---
title: "R Notebook"
output: html_notebook
---

source(file = "~/R Fat Integer/WGCNA.R")

# Correlation between fatty acids ACP dimensions and modules eigengenes:
```{r}
load(file = "./rdata/clustering/clust_sh.RData")
wgcna_data = clust_sh

load(file = "./rdata/pca/pca_fa.RData")
FA_PCA = as.data.frame(res.pca$ind$coord[order(as.numeric(rownames(res.pca$ind$coord))),], row.names = rownames(wgcna_data$MEs))

load(file = "./rdata/pca/pca_met.RData")
met_PCA =  as.data.frame(res.pca$ind$coord[order(as.numeric(rownames(res.pca$ind$coord))),],row.names = rownames(wgcna_data$MEs) )
```


```{r}
significatifmodule = wgcna_data
significatifmodule$MEs = significatifmodule$MEs[c(
                                                  "MEbrown", 
                                                  "MEdarkgreen", 
                                                  "MEdarkolivegreen", 
                                                  "MEdarkorange",
                                                  "MEdarkred",
                                                  "MEgreen", 
                                                  "MElightcyan",
                                                  "MEroyalblue",
                                                  "MEsaddlebrown", 
                                                  "MEskyblue",
                                                  "MEsteelblue",
                                                  "MEviolet",
                                                  "MEwhite"
                                                  )]
```

# Metabolomic data
## ALL MODULES 
```{r}
nGenes = length(wgcna_data$probes);
nSamples = nrow(met_PCA);

MEs = wgcna_data$MEs
modulePCACor = cor(MEs, met_PCA, use = "p");
modulePCAPvalue = corPvalueStudent(modulePCACor, nSamples);

pdf("./figures/pca_met_heatmap.pdf", width = 8, height = 15)

# Will display correlations and their p-values
textMatrixPCA =  paste(signif(modulePCACor, 2), "\n(",
                    signif(modulePCAPvalue, 1), ")", sep = "");

par(mar = c(5, 9, 3, 3));

# Display the correlation values within a heatmap plot
hm = labeledHeatmap(Matrix = modulePCACor,
                    xLabels = colnames(modulePCACor),
                    yLabels = names(MEs),
                    ySymbols = names(MEs),
                    colorLabels = FALSE,
                    colors = blueWhiteRed(50),
                    naColor = "white",
                    xLabelsAngle = 0,
                    xLabelsPosition = "top",
                    xLabelsAdj = 0.5,
                    textMatrix = textMatrixPCA,
                    setStdMargins = FALSE,
                    cex.text = 1,
                    zlim = c(-1,1),
                    showRows = NULL,
                    font.lab.x = 1.5)
dev.off()

```
![Heatmap met all modules](./figures/pca_met_heatmap.pdf){width=60%}

## SIGNIFICANT MODULES FOR FEED EFFICIENCY
```{r}

nGenes = length(significatifmodule$probes);
nSamples = nrow(met_PCA);

MEs = significatifmodule$MEs
modulePCACor = cor(MEs, met_PCA, use = "p");
modulePCAPvalue = corPvalueStudent(modulePCACor, nSamples);

pdf("./figures/pca_met_heatmap_sign.pdf", width = 8, height = 8)

# Will display correlations and their p-values
textMatrixPCA =  paste(signif(modulePCACor, 2), "\n(",
                       signif(modulePCAPvalue, 1), ")", sep = "");

par(mar = c(5, 9, 3, 3));

# Display the correlation values within a heatmap plot
hm = labeledHeatmap(Matrix = modulePCACor,
                    xLabels = colnames(modulePCACor),
                    yLabels = names(MEs),
                    ySymbols = names(MEs),
                    colorLabels = FALSE,
                    colors = blueWhiteRed(50),
                    naColor = "white",
                    xLabelsAngle = 0,
                    xLabelsPosition = "top",
                    xLabelsAdj = 0.5,
                    textMatrix = textMatrixPCA,
                    setStdMargins = FALSE,
                    cex.text = 1,
                    zlim = c(-1,1),
                    showRows = NULL,
                    font.lab.x = 1.5)
dev.off()

```
![Heatmap met significant modules](./figures/pca_met_heatmap_sign.pdf){width=60%}

# Fatty acids data
## ALL MODULES 
```{r}
nGenes = length(wgcna_data$probes);
nSamples = nrow(FA_PCA);

MEs = wgcna_data$MEs
modulePCACor = cor(MEs, FA_PCA, use = "p");
modulePCAPvalue = corPvalueStudent(modulePCACor, nSamples);

pdf("./figures/pca_FA_heatmap.pdf", width = 10, height = 15)

# Will display correlations and their p-values
textMatrixPCA =  paste(signif(modulePCACor, 2), "\n(",
                    signif(modulePCAPvalue, 1), ")", sep = "");

par(mar = c(5, 9, 3, 3));

# Display the correlation values within a heatmap plot
hm = labeledHeatmap(Matrix = modulePCACor,
                    xLabels = colnames(modulePCACor),
                    yLabels = names(MEs),
                    ySymbols = names(MEs),
                    colorLabels = FALSE,
                    colors = blueWhiteRed(50),
                    naColor = "white",
                    xLabelsAngle = 0,
                    xLabelsPosition = "top",
                    xLabelsAdj = 0.5,
                    textMatrix = textMatrixPCA,
                    setStdMargins = FALSE,
                    cex.text = 1,
                    zlim = c(-1,1),
                    showRows = NULL,
                    font.lab.x = 1.5)
dev.off()

```
![Heatmap met all modules](./figures/pca_FA_heatmap.pdf){width=60%}

## SIGNIFICANT MODULES FOR FEED EFFICIENCY
```{r}

nGenes = length(significatifmodule$probes);
nSamples = nrow(FA_PCA);

MEs = significatifmodule$MEs
modulePCACor = cor(MEs, FA_PCA, use = "p");
modulePCAPvalue = corPvalueStudent(modulePCACor, nSamples);

pdf("./figures/pca_FA_heatmap_sign.pdf", width = 8, height = 8)

# Will display correlations and their p-values
textMatrixPCA =  paste(signif(modulePCACor, 2), "\n(",
                       signif(modulePCAPvalue, 1), ")", sep = "");

par(mar = c(5, 9, 3, 3));

# Display the correlation values within a heatmap plot
hm = labeledHeatmap(Matrix = modulePCACor,
                    xLabels = colnames(modulePCACor),
                    yLabels = names(MEs),
                    ySymbols = names(MEs),
                    colorLabels = FALSE,
                    colors = blueWhiteRed(50),
                    naColor = "white",
                    xLabelsAngle = 0,
                    xLabelsPosition = "top",
                    xLabelsAdj = 0.5,
                    textMatrix = textMatrixPCA,
                    setStdMargins = FALSE,
                    cex.text = 1,
                    zlim = c(-1,1),
                    showRows = NULL,
                    font.lab.x = 1.5)
dev.off()

```
![Heatmap met significant modules](./figures/pca_FA_heatmap_sign.pdf){width=60%}
