---
title: "R Notebook"
output: html_notebook
---


```{r}
setwd("~/R Fat Integer/data");

if (!require("dplyr")) install.packages("dplyr")   
if (!require("do")) install.packages("do")   
if (!require("stringr")) install.packages("stringr")  
if (!require("ggplot2")) install.packages("ggplot2")  
```

```{r}
source(file = "./WGCNA.R")
```
```{r}
load(file = "./rdata/datainput/datainput.RData")
```


```{r}
PlotSoftThresholdingPower(datExprFoldChange, "signed hybrid")
```
$\beta$ is a non-dichotomic soft thresholding that allows to evaluate connection between probes without losing the continuous character of the co-expression. Low correlations are better masked with high $\beta$ values. We set $\beta$ to $6$ according to the criterion of the “approximate scale free topology”.

WGCNA clustering : 
```{r}
clust_sh = Clustering.WGCNA(dataExpr = data, TOMRData = NULL, TOM = NULL, minModuleSize = 25, softPower = 6, networkType = "signed hybrid", deepSplit = 2)
```

```{r}
table(rownames(clust_sh$MEs) == rownames(datTraits))
```
Composition of the modules (probes, genes and unique genes):
```{r}
list.probes = ModuleComposition.Probes(clustering = clust_sh, cat = "blood")
list.genes = ModuleComposition.Genes(clustering = clust_sh, cat = "blood", annotation = annotation)
list.genes.uniq = lapply(ModuleComposition.Genes(clustering = clust_sh, cat = "blood", annotation = annotation),unique_no.NA)
```

```{r}
size.probes = as.data.frame(table(clust_sh$dynamicColors))
size.genes = as.data.frame(t(as.data.frame(lapply(list.genes.uniq[], length))))

nGenes = length(clust_sh$probes);
nSamples = nrow(datTraits);

MEs = clust_sh$MEs

moduleTraitCor = cor(MEs, datTraits, use = "p");
moduleTraitPvalue = corPvalueStudent(moduleTraitCor, nSamples);

size.probes$Var1 = paste("ME", size.probes$Var1, sep="")
table(size.probes$Var1 == rownames(moduleTraitCor))

size.genes$Var1 = size.probes$Var1
table(size.probes$Var1 == rownames(moduleTraitCor))

moduleTraitCor2 = cbind(rep(c(NA), each=34),moduleTraitCor)
moduleTraitPvalue2 = cbind(size.genes$V1, moduleTraitPvalue)

moduleTraitCor3 = cbind(rep(c(NA), each=34),moduleTraitCor2)
moduleTraitPvalue3 = cbind(size.probes$Freq, moduleTraitPvalue2)
```

Display the correlation values within a heatmap plot (Modules-Traits relationships):
```{r}
# Will display correlations and their p-values
textMatrix =  paste(signif(moduleTraitCor, 2), "\n(",
                    signif(moduleTraitPvalue, 1), ")", sep = "");
textMatrix2 = c(size.genes$V1, textMatrix)
textMatrix2 = c(size.probes$Freq, textMatrix2)


label = str_sub(string = names(MEs),start = 3, end =-1)

colnames(moduleTraitCor3) <- c("probes", 'unique\n genes', "num", "line", "diet", "age", "weight", "ADG", "ADFI", "FCR", "%loin", "%backfat" )

# Display the correlation values within a heatmap plot

pdf("./figures/clust_heatmap.pdf", width = 10, height = 15)
par(mar = c(5, 9, 3, 3));

labeledHeatmap(Matrix = moduleTraitCor3,
               xLabels = colnames(moduleTraitCor3),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               naColor = "white",
               xLabelsAngle = 0,
               xLabelsPosition = "top",
               xLabelsAdj = 0.5,
               textMatrix = textMatrix2,
               setStdMargins = FALSE,
               cex.text = 1,
               zlim = c(-1,1),
               showCols = c(1, 2, 7 ,8, 9, 10, 11, 12),
               font.lab.x = 1.5,
               height = 20)
dev.off()
```

![clust heatmap](./figures/clust_heatmap.pdf){width=60%}



Distribution of module size :
```{r}
ggplot(size.probes, aes(Var1,Freq))+theme(axis.text.x = element_text(angle = 90)) + geom_bar(stat='identity', fill = "grey70")+geom_text(aes(label = Freq, y = Freq), size=4) + xlab("Modules") + ylab("Size (number of probes)")
```


```{r}
blood_highly_significant = c("violet", "darkred", "lightcyan")
blood_very_significant = c()
blood_significant = c("royalblue", "white", "darkorange")
blood_trend = c("darkolivegreen", "steelblue")
```

### Hierarchical clustering of module eigengenes.
The modules that were found highly correlated with feed conversion ratio (FCR) are enlightened (*** $P \leq 0.001$, * $P \leq 0.05$, and a dagger $0.05 < P \leq 0.10$).
```{r}
METree = clust_sh$METree
for (i in c(1:length(METree$labels))){
  if (is.element(substring(METree$labels[i],3),blood_highly_significant)){
      METree$labels[i] = paste("***", METree$labels[i])
  }
  if (is.element(substring(METree$labels[i],3),blood_very_significant)){
    METree$labels[i] = paste("**", METree$labels[i])
  }
  if (is.element(substring(METree$labels[i],3),blood_significant)){
    METree$labels[i] = paste("*", METree$labels[i])
  }
  if (is.element(substring(METree$labels[i],3),blood_trend)){
    METree$labels[i] = paste("†", METree$labels[i])
  }
}

sizeGrWindow(7, 6)
plot_METree = plot(METree, xlab = "", main ="", sub = "")

```


### Calculate ModuleMembership and the Gene Significance for FCR:
```{r}
dataExpr = datExprFoldChange

intramod = intramodularConnectivity.fromExpr(dataExpr, colors = clust_sh$dynamicColors, networkType = "signed hybrid")
rownames(intramod) = colnames(dataExpr)
intramod$dynamicColors = clust_sh$dynamicColors

# Define variable FCR containing the weight column of datTrait
FCR = as.data.frame(datTraits$FCR);
names(FCR) = "FCR"

# names (colors) of the modules
modNames = substring(names(clust_sh$MEs), 3)

geneModuleMembership = as.data.frame(cor(dataExpr, clust_sh$MEs, use = "p"));
MMPvalue = as.data.frame(corPvalueStudent(as.matrix(geneModuleMembership), nSamples));

names(geneModuleMembership) = paste("MM", modNames, sep="");
names(MMPvalue) = paste("p.MM", modNames, sep="");

geneModuleMembership$ProbeName = rownames(geneModuleMembership)
MMPvalue$ProbeName = rownames(MMPvalue)
MM = right_join(geneModuleMembership, MMPvalue, by=c("ProbeName"))
table(MM$ProbeName == clust_sh$probes)
MM$modules = clust_sh$dynamicColors

MM = right_join(annotation, MM, by=c("ProbeName"))

geneTraitSignificance = as.data.frame(cor(dataExpr, FCR, use = "p"));
GSPvalue = as.data.frame(corPvalueStudent(as.matrix(geneTraitSignificance), nSamples));

names(geneTraitSignificance) = paste("GS.", names(FCR), sep="");
names(GSPvalue) = paste("p.GS.", names(FCR), sep="");

geneTraitSignificance$ProbeName = rownames(geneTraitSignificance)
geneTraitSignificance$GSPvalue = GSPvalue$p.GS.FCR
table(geneTraitSignificance$ProbeName == clust_sh$probes)
geneTraitSignificance$modules = clust_sh$dynamicColors

geneTraitSignificance = right_join(annotation,geneTraitSignificance,by=c("ProbeName"))
geneTraitSignificance$absGS = abs(geneTraitSignificance$GS.FCR)
```
Top 10 Gene Significance for each module :
```{r}
geneTraitSignificance %>% group_by(modules) %>% top_n(10, absGS) %>% arrange(modules)
```




#### Save clustering and composition :
```{r}
save(clust_sh, list.genes, list.genes.uniq, list.probes, file = "./rdata/clustering/clust_sh.RData" )
```


